---
title: "Identify and Visualize Potential Swaps in Longitudinal Weight Data"
author: "Your Name"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Identify and Visualize Potential Swaps in Longitudinal Weight Data}
  %\VignetteEngine{knitr::rmarkdown}
---

# Identify and Visualize Potential Swaps in Longitudinal Weight Data

In this vignette, we will walk you through how to use the `identify_and_plot_swaps` function from the `cleanR` package. This function identifies potential swaps in longitudinal weight data where the weight in an earlier timepoint exceeds the weight in a later timepoint. The function will flag these rows and generate a plot to help visualize the flagged data.

## Function Overview

The `identify_and_plot_swaps` function:
- Flags rows where the weight in an earlier timepoint exceeds the weight in a later timepoint (potential swaps).
- Highlights flagged points in a line graph for easy visualization.
- Returns the dataset with flagged rows, the line graph, and the path to the saved plot.

### Parameters:
- `data`: A data frame containing the dataset.
- `weight_columns`: A character vector specifying the column names containing weight data for different timepoints.
- `key_column`: A character string specifying the column name used to identify rows (e.g., "pot_id").
- `plot_filename`: A character string specifying the filename for saving the plot (e.g., "swaps_plot.png").

### Output:
The function returns a list containing:
- `swaps`: A data frame of rows flagged as potential swaps.
- `plot`: A ggplot object visualizing the weights and flagged potential swaps.
- `saved_plot`: The path to the saved image file.

## Example Usage

### Loading the Data

Let's start by loading the sample dataset:

```{r}
library(cleanR)

# Example dataset
plants_clean <- read.csv("/cloud/project/cleanR/data-raw/plants_test.csv")

# Preview the data
head(plants_clean)
```

Identify and Visualize Potential Swaps
Now we can use the identify_and_plot_swaps function to identify potential swaps in the weight data:
```{r}
# Call the function to identify and visualize swaps
swaps_output <- identify_and_plot_swaps(
  data = plants_clean,
  weight_columns = c("X10_6_weight", "X10_9_weight"), 
  key_column = "pot_id", 
  plot_filename = "swaps_plot.png"
)

# View the flagged swaps
print(swaps_output$swaps)
```
Understanding the Output
The function will return a list that includes:

Swaps: The rows with flagged potential swaps, where earlier timepoints have greater weights than later ones.
Plot: A line plot showing the longitudinal weight data, with flagged points marked.
Saved Plot: The filename where the plot was saved.
Example Plot
The function generates a plot where each line represents the weight data for a specific plant (identified by pot_id). Points flagged as potential swaps will appear as crosses, indicating that the weight in an earlier timepoint is greater than the weight in a later timepoint.

Conclusion
The identify_and_plot_swaps function is a useful tool for detecting and visualizing potential errors in longitudinal weight data. By flagging potential swaps and generating an easy-to-read graph, it helps ensure the integrity of the data and facilitates error detection.
