---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# cleanR

<!-- badges: start -->
<!-- badges: end -->

The goal of cleanR is to equip users with functions that can make identifying potential errors in datasets a little bit easier. 

## Description

The `cleanR` package provides tools for cleaning and processing data, via the following functions: 

1. Coercing non-numeric values to numeric: 
This function takes a column that a user assumes should have numeric inputs, but which may have non-numeric values such as word (example: "dead"). Regular coercion can turn values into NAs, which can disrupt any operations you might want to perform on the data down the line such as mathematical operations. This function coerces the column to numeric, but is also set to deal with special cases such as "dead". It alerts the user of where there were non-numeric values. 

2. Identifying typos:
This function analyzes a character column to identify potential typos by using string length and distance checks against the most common string in that column (which it uses as the non-typo string). 

3. Identifying potential swaps via visualizing data: This function is designed to compare weight measurements between separate timepoints to identify potential swaps that might have occured in the data, where the weight for one row was swapped for the weight for another row. It specifically works in datasets where the weight increases or is expected to increase over time. It also provides user with a visualization of the results so that it is easier to see the potential swap that took place, rather than looking at huge swaths of numbers. 

4. Identiying invalid lines (specific to my example dataset):
Additionally, this package contains a 4th function that is particularly suited to example dataset provided (plants_clean). This is a cut of data collected by me for a future project. In this dataset, plants are tracked by weight over months (only two days are included in the cut of data used for the example), and have been divided into two lines that will eventually receive different watering treatments to achieve conditions of drought (treatment:"d") or control (treatment: "w"). The plants used are the result of artificial selection done to create families that are either self-incompatible -> SI (the plant cannot reproduce with itself), self-compatible -> SC(the plant can reproduce with itself), or control -> C. This will become important later on in my project when we begin to do experiments on them to understand how their reproductive abilities are responding to the drought treatment, so it is important that these lines are correctly labelled in the dataset. Multiple people in my lab work with these lines as well, so this function is applicable to other datasets. Because of that, I wanted to create a function that addressed this specific feature of my data. This is function #4 and it identifies invalid lines in the data, prints out the result to the user and suggests the line IDs that the should be instead. 

## Installation

Here's how you can install the cleanR package: 

``` r
# Install the following packages:
install.packages(c("devtools", "readr", "stringdist"))



# Install cleanR from GitHub
devtools::install_github("niafaithlewis/cleanR/cleanR")
```

## Example 1

Here is an example of the invalid lines function in action: 

```{r}
### Load the required packages
library(cleanR)

### Load data
data("plants_clean")

head(plants_clean)

invalid_lines <- detect_invalid_lines(plants_clean)

```



## Example 2

This is an example of how to use the find_swaps function to identify and visualize potential swaps in the data:

Note: this function will drop NA values from the data, while converting columns to numeric. 

```{r example}
### Load libraries 
library(ggplot2)
library(tidyr)
library(cleanR)

data("plants_clean")


### Apply the identify_and_plot_swaps function
swaps_output <- identify_and_plot_swaps(
  data = plants_clean,
  weight_columns = c("X10_6_weight", "X10_9_weight"),
  key_column = "pot_id",  
  plot_filename = "swaps_plot.png",  
  weight_diff_threshold = 100  
)


### View the flagged swaps
print(swaps_output$swaps)
```



![The plot will be saved under "inst/extdata/swaps_plot"! This plot shows the potential swaps as thicker lines than the rest of the graph to highlight them to the user. The IDs that have been flagged aas potential swaps (Flagged = TRUE) also have a distinct marker than those that are not likely to be swaps.](inst/extdata/swaps_plot.png)

Find more detailed examples in the vignettes folder. The vignettes provide detailed explanations of the functions and show examples of how to use each one, and what results you can expect. 

You can access the vignettes in the vignettes folder or by viewing them directly: 

1. Coercing non-numeric values to numeric 
```{r}
vignette("coerce_to_numeric")
```

2. Identifying typos
```{r}
vignette("identifying_typos")
```
3. Identifying potential swaps via visualizing data
```{r}
vignette("find_swap")
```

4. Identifying invalid lines
```{r}
vignette("invalid_lines")
```

