---
title: "Detecting Invalid 'Line' Values"
author: "Your Name"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Detecting Invalid 'Line' Values}
  %\VignetteEngine{knitr::rmarkdown}
---

# Detecting Invalid 'Line' Values

In this vignette, we will demonstrate how to use the `detect_invalid_lines` function from the `cleanR` package to identify invalid `line` values in your dataset. This function helps flag rows where the `line` column does not match the expected patterns, such as `"C1"`, `"SC1"`, or `"SI1"`. It also suggests potential corrections for these invalid entries.

## Function Overview

The function `detect_invalid_lines` checks the `line` column of your dataset for values that do not match one of the valid patterns. If any invalid values are found, the function:
- Identifies the invalid rows.
- Suggests possible corrections based on string distance (e.g., using the Jaro-Winkler method).

### Parameters:
- `data`: A data frame that contains the dataset with a `line` column.
- The function will return a data frame with the invalid rows, including suggested corrections for each invalid `line` value.

## Example Usage

### Loading the Data

Let’s first load a sample dataset and inspect the `line` column:

```{r}
library(cleanR)

# Example dataset
mousedata <- data.frame(
  treatment = c("d", "d", "d", "wet", "d", "w"),
  pot_id = c(1, 2, 3, 4, 5, 6),
  line = c("C1", "SXX1", "SC1", "SI1", "C1", "C1"),
  mom = c(86, 225, 231, 672, 2, 27),
  dad = c(59, 258, 245, 697, 44, 100),
  tray_id = c(1, 10, 1, 1, 1, 1),
  dry_weight = c(165, 168, 158, 170, 172, 160)
)

# Preview the data
head(mousedata)
```

Now, applying the function to the dataset 
```{r}
# Detect invalid 'line' values
invalid_lines <- detect_invalid_lines(mousedata)

# Print the result
print(invalid_lines)
```

Explanation of Output 